---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}

activity_data=read.csv("activity.csv", header = T)
head(activity_data)
str(activity_data)
dim(activity_data)[1]

# remove NA in data
data1 <- activity_data[ with (activity_data, { !(is.na(steps)) } ), ]
head(data1)
str(data1)
table(data1$interval)
```


## What is mean total number of steps taken per day?
```{r}
#Calculate the total number of steps taken per day
library(dplyr)

by_date<- group_by(data1, date)
total_number<- summarise(by_date,total_step = sum(steps))


## histogram of of total step by date
hist(total_number$total_step, main="Histogram Total Steps per Day", xlab="# Steps", col="red") # plot a histogram  


##Calculate and report the mean and median of the total number of steps taken per day
summary(total_number$total_step)
```



## What is the average daily activity pattern?
```{r}
# preprocessing data for plot
steps_by_interval <- aggregate(steps ~ interval, data1, mean)

# create a time series plot 
plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
     main="Average number of steps over all days", xlab="Interval", 
     ylab="Average number of steps")

###finding the maximum interval
#which.max indicate for us the position. 
steps_by_interval[which.max(steps_by_interval$steps),]
```



## Imputing missing values
```{r}

#Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
table(is.na(activity_data$steps))


#imputing all those NA step row in the orignal data with the mean steps from each 5 mins intevral timing. 
new_data=activity_data

for (i in 1: dim(new_data)[1]){
  if(is.na(new_data$steps[i])){
    interval_vaue=new_data$interval[i]
    new_data$newsteps[i]=steps_by_interval$steps[steps_by_interval$interval==interval_vaue]
  } else new_data$newsteps[i]=new_data$steps[i]
}

#checking the impute data
head(new_data[is.na(new_data$steps),])

#histogram for the new_data
by_date_new<- group_by(new_data, date)
total_numbe_newr<- summarise(by_date_new,total_step_new = sum(newsteps))


## histogram of of total step by date
hist(total_numbe_newr$total_step_new, main="Histogram Total Steps per Day", xlab="# Steps", col="red") # plot a histogram  

##Calculate and report the new_mean and new_ median of the total number of steps taken per day
summary(total_numbe_newr$total_step_new)[3]

#comparing the 2 mean and median

mean1_vs_mean2=c(summary(total_number$total_step)[4],summary(total_numbe_newr$total_step_new)[4])
median1_vs_median2=c(summary(total_number$total_step)[3],summary(total_numbe_newr$total_step_new)[3])

```



## Are there differences in activity patterns between weekdays and weekends?

#creating a new factor for the new dataset.
new_data$new_weekday=weekdays(new_data$date)
head(new_data)

#as weekdays function only allow monday to sunday, we need to categeroise them into 2
new_data$new_type=ifelse(((new_data$new_weekday=="Sunday")|(new_data$new_weekday=="Saturday")),"Weekends","weekdays")
head(new_data)

#exporting the data
write.table(new_data, "~/GitHub/RepData_PeerAssessment1/mydata.txt", sep="\t")
```{r}
head(new_data)
new_data1=read.table("mydata.txt", , sep="\t")

steps_by_interval_1=aggregate(newsteps~interval+new_type,new_data1,mean)
library(ggplot2)
# creat a plot
qplot(interval, 
      newsteps, 
      data = steps_by_interval_1, 
      type = 'l', 
      geom=c("line"),
      xlab = "Interval", 
      ylab = "Number of steps", 
      main = "") +
  facet_wrap(~ new_type, ncol = 1)


```
